### 2012-05-04 paper/paper.tex
moar writeup

\documentclass{sig-alternate}

\usepackage{xspace}
\newcommand\Motmot{\textsc{Motmot}\xspace}

\begin{document}

\CopyrightYear{2012}

\title{Motmot: A Distributed Chat Protocol}

\numberofauthors{5}
%
\author{
  \alignauthor
  Carl Jackson \\
    \affaddr{Harvard University} \\
    \email{cjackson@college.harvard.edu}
  %
  \alignauthor
  Max Wang \\
    \affaddr{Harvard University} \\
    \email{max.wang@college.harvard.edu}
  %
  \alignauthor
  EJ Bensing \\
    \affaddr{Harvard University} \\
    \email{ebensing@college.harvard.edu}
  %
  \and
  \alignauthor
  Jeff Atwood \\
    \affaddr{Harvard University} \\
    \email{jatwood@college.harvard.edu}
  %
  \alignauthor
  Julie Hansbrough \\
    \affaddr{Harvard University} \\
    \email{jhansbrough@college.harvard.edu}
}

\maketitle

\begin{abstract}

\Motmot is a new distributed text chat protocol based on Paxos. Unlike previous
chat protocols like XMPP, OSCAR, and IRC, \Motmot provides guaranteed message
delivery, total message ordering, end-to-end encryption, peer authentication,
and message deniability.
% TODO: server federates, etc.
% TODO: no xml! It's a feature
% TODO: write more

\end{abstract}

\section{Introduction}

\section{libmotmot}

Guaranteed delivery and total message ordering are perhaps the most fundamental
chat features provided by \Motmot.  In order to obtain these properties, the
clients participating in a chat must reach a consensus concerning the reception
and order of chat events.  \Motmot assumes a benign failure model and uses a
variant of the Paxos consensus protocol \cite{paxos}, which is the core of its
client library \texttt{libmotmot}, in order to provide these guarantees.

\subsection{Design}

\subsection{Implementation}

\subsection{Testing}

\subsection{Lessons Learned}

\section{Discovery Server}

The discovery server performs several tasks in the system: authentication and
identity federation; presence management; and connection brokering.

\subsection{Identity}

Much like other federated protocols like XMPP and email, identity handles are
based upon DNS. Every domain name with a SRV record can act as a \Motmot
identity realm, and give out email-like addresses (an \verb`addr-spec` as
defined in RFC 2822) for \Motmot communication.

To grant a user an identity on a realm, the user first presents a username and
password to the server, after which the server delegates authority for that name
to the client by signing that client's key. This key-signing process is the
method by which peers can authenticate securely to each other without a
centralized broker.

\subsection{Presence}

In chat protocols, presence is the system by which users can specify a list of
buddies to publish their online status to. The \Motmot discovery server provides
a presence lookup service, by recording the status of all of the clients
connected to it. Servers then push presence updates to all of the other servers
on any user's buddy list, which in turn push the notification to the relevant
clients.

This subsystem system does not need to be robust or reliable, since all pushed
state can be recreated from other sources, and since asynchronously pushed
presence notifications do not need to be perfectly reliable to ensure a usable
user interface.

\subsection{Brokering}

Finally, the server must perform message brokering. For clients behind firewalls
or routers which perform network address translation (NAT), direct connections
are not possible. In such a case, the discovery server might be forced to assist
in connecting the clients.

Technologies like STUN, TURN, and ICE provide implementations of ``NAT poking,''
and while mechanisms for interfacing with such servers were considered in the
design of the server, performing brokering itself was outside the scope of this
version of the protocol.

\subsection{Implementation}

The \Motmot discovery server was implemented in Python using several open source
libraries. The server itself is a custom RPC-like reactor/dispatcher written
% TODO: cite gevent
using gevent, a coroutine-based networking library. For our communication
% TODO: cite msgpack
protocol, we use msgpack to serialize messages. This allows for easy
cross-language communication with clients and other servers, since
implementations are available for most popular languages.

We found that these two libraries provided good abstractions for the task of
implementing a server. Gevent's coroutines provided an extremely useful
concurrency model: they were executed in a single thread, so synchronization
around global objects was not necessary, while avoiding the complexity (and
unnecessary indentation) of a callback-based system. Msgpack also provided an
excellent abstraction, allowing the server to assume that arbitrary data
structures could be written to and read from the wire. The ad-hoc nature of the
format also helped speed development of the server and protocol.

% TODO: cite pyopenssl
Finally, we used openssl (and the pyOpenSSL library in particular) to perform
certificate operations.

\subsection{Lessons Learned}

% TODO: EJ?

\section{Pidgin Plugin}

To allow our protocol to be utilized via Pidgin, we needed our plugin to interface with two major aspects of the protocol: the discovery server, and the libmotmot library.

\subsection{Plugin Interactions with Discovery Server}

\subsection{Plugin Interactions with Libmotmot}
There were a few major difficulties that had to be addressed when calling libmotmot within the plugin:

\begin{itemize}
\item libpurple makes a distinction between "IMs" (two-person chats) and "chats" (three or more people); libmotmot does not.  libpurple takes in different arguments for its IM-related and its chat-related functions

\item The main Pidgin event loop cannot be modified from within a plugin (without patching libpurple itself).  Instead, callbacks are made to plugin functions from within the main event loop, so we had to find a way to call libmotmot functions correctly (i.e. still honoring the correct control flow from libmotmot's perspective) and consistently within the bounds of these plugin functions

\item libmotmot and libpurple use different handles to differentiate between chats and users, and translating between the two handles proved challenging, especially given the complexity of the information associated with libpurple chats (e.g. a single chat may have an id, a room name, PurpleConnection and PurpleConvChat objects...)

\end{itemize}

The bulk of the integration work was writing callbacks for the motmot\_init() function of libmotmot.  Once these callbacks are written correctly, libmotmot handles most of the work of starting, sending, and coordinating chats automatically.  These callbacks were, in almost all instances, responsible for calling libpurple functions, which handle the UI aspect of the associated libmotmot calls.  In order to pass relevant information to our callbacks, we created a MotmotInfo struct, a large data struct that is used differently in different callbacks to get all the necessary info to the libpurple calls.

We faced a challenge when writing the enter() function.  This function is a callback responsible for allowing a user to enter a chat.  Due to the structure of libmotmot, the only information passed into enter() as an argument is an internal handle, and we need create some sort of external handle.  We solved this by creating a global variable, GLOBAL\_ACCOUNT, which represents the account that a user first logs in under, and we associate our external handle with this GLOBAL\_ACCOUNT.  Though this ``works'' if a user only signs in with on eaccount, this breaks when a user decides to sign in with multiple accounts at once (something that libpurple allows), and should be addressed in future versions.

\section{Conclusions}

Wheeee.

\section{Acknowledgments}

\thebibliography{2}{
  \bibitem{paxlive}
    T. Chandra, R. Griesemer, and J. Redstone.
    Paxos made live---an engineering perspective.
    In \emph{Proc. 26th Symp. on Principles of Distributed Computing}, pp. 398-407, Aug. 2007.

  \bibitem{paxos}
    L. Lamport.
    The part-time parliament.
    \emph{ACM Transactions on Computer Systems}, 16(2):133-169, 1998.

  \bibitem{paxsimp}
    L. Lamport.
    Paxos made simple.
    \emph{ACM SIGACT News 32}, 4:18-25, Dec. 2001.

  \bibitem{rfc}
    P. Resnick, ed.
    Internet Message Format, RFC 2822.
    April 2001.
}

\end{document}