### 2012-11-12 6.858/proposal/proposal.tex
Stub out project proposal

\documentclass[12pt]{article}

\usepackage{geometry, xspace}
\usepackage[parfill]{parskip}
\geometry{letterpaper}

\newcommand{\trill}{\textsc{Trill}\xspace}

\title{Trill: Secure firewall-traversing networking}
\author{Carl Jackson and Max Wang}
\date{November 5, 2012}

\begin{document}

\maketitle

\section*{Overview}
We propose to design and implement \trill, a networking abstraction that
provides secure communication between federated identities, either of which
could be behind a firewall.

\section*{Security}
% TODO

\section*{Federated Identity}
% TODO

\section*{Firewall Traversal}
% TODO

\section*{Networking Abstractions}
% TODO

\end{document}