### 2012-12-14 6.858/paper/paper.tex
Make the proposal past tense
This is a good first approximation of a final paper.

\documentclass[12pt]{article}

\usepackage{fancyhdr, geometry, hyperref, xspace}
\usepackage[parfill]{parskip}

\geometry{letterpaper,text={7in,9in},centering}
\pagestyle{fancy}

\fancyhead[RO]{Carl Jackson and Max Wang}
\fancyhead[LO]{\url{http://github.com/motmotchat/trill}}

\newcommand{\Trill}{\textsc{Trill}\xspace}

\title{\Trill: Secure firewall-traversing networking}
\author{Carl Jackson and Max Wang}
\date{December 14, 2012}

\makeatletter
\renewcommand{\maketitle}{%
  \thispagestyle{plain}%
  \begin{center}%
    {\LARGE \@title \par}%
    {\large \@author \par}%
    {\large \@date \par}%
  \end{center}%
}
\makeatother

\begin{document}

\maketitle

\section*{Overview}
We implemented \Trill, a networking abstraction that provides secure
communication between federated identities, either of which could be behind a
firewall.

\section*{Security}
The security of our system is based on standard cryptographic infrastructure and
libraries. By ``identities,'' we mean public key pairs, the private half of
which is uniquely known by an individual. This identity can be given one or more
names by means of X.509 certificates, and standard mechanisms like TLS allow
owners of names to securely establish a communication channel and mutually
verify the other's identity. Due to implementation restrictions in our target
environment, we used DTLS, a younger protocol that uses UDP as its transport
layer; specifically, we used the implementation of DTLS provided by the open
source GnuTLS library.

\section*{Federated Identity}
As with other federated protocols like XMPP and email, \Trill identities are
based on DNS.  Every domain name with a SRV record can act as a \Trill identity
realm, supporting email address-like identities (an \verb!addr-spec!  as defined
in RFC 2822).

Each \Trill domain hosts a federation server, which is able to delegate
authority for identities to various clients.  Any client may present the server
with an identity and a password for that identity.  If the client's
authentication succeeds, the server signs the client's key and returns the
resultant certificate.  Using these certificates, clients can then authenticate
securely to one another without further use of the server.

\section*{Firewall Traversal}
Due to restrictions on many firewalls and NATs in place on the Internet, it is
impossible for arbitrary clients on the internet to establish peer-to-peer
communication. There are techniques for firewall/NAT traversal which involve
``tricking'' the networking layer into establishing a connection. We implemented
something similar to \texttt{chownat} (\url{http://samy.pl/chownat/}) in \Trill,
which appears to be a standard way to do this.

\end{document}